---
title: "Customer Churn Analysis"
output: 
    html_document:
        smart: false
---

Loading the necessary pacakges for this analysis: 

```{r messages=FALSE}

library(dplyr)
library(psych)
library(tidyr)
library(ggplot2)
library(rpart)
library(rpart.plot)
library(caTools)
library(randomForest)
library(knitr)
library(rmarkdown)
library(markdown)
library(kableExtra)


```

```{r echo=FALSE} 
options(knitr.table.format = "html")
df  = read.csv("D:\\Emman\\Documents\\GitHub\\TelcoChurn\\ChurnAnalysis\\ChurnAnalysis\\datasource\\customerchurn.csv", na.strings = c("", "NA"))

kable(cbind(head(df), head(df))) %>%
     kable_styling(bootstrap_options = "striped",
                full_width = F) %>%
             scroll_box(width='100%')

kable(describe(df), format = "html") %>%
    kable_styling(bootstrap_options = "striped",
                full_width = F) %>%
    scroll_box(width='100%')
```


This is an R Markdown is document. Markdown is a simple formatting syntax for authoring HTML, PDF, and Microsoft Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you are click the **R Tools | Publish | Preview** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
